---
title: '`Course Project 2` Reproducible Research'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
params:
  author: "Anderson H Uyekita"
  course: "Reproducible Research"
  course_url: "https://www.coursera.org/learn/reproducible-research"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  course_start: !r lubridate::ymd("2022/06/24")
  course_finish: !r lubridate::today()
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project-2_reproducible-research'
  project_repo: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2'
  codebook: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
library(shiny)
```


`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s)", "./instructions.md")}else{sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s){target='_blank' rel='noopener'}", params$instructions)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4C4; README: [README.md](%s)", "./README.md")}else{sprintf(fmt = "* &#x1F4C4; README: [README.md](%s){target='_blank' rel='noopener'}", params$project_repo)}`

--------------------------------------------------------------------------------

## Codebook




### 1. Requirements

```{r,echo=TRUE,message=FALSE,warning=FALSE,include=FALSE}
library(tidyverse)
library(pryr)
library(kableExtra)
library(magrittr)
```

### 2. Scripts

#### 2.1. README

### 3. Output and Input Details

Course Project 2 has only one input dataset from NOAA.

#### 3.1. Inputs

```{r loading-storm-data,echo=TRUE,cache=TRUE}
# Loading raw data.
raw_data <- read.csv(file = "./data/repdata_data_StormData.csv.bz2")

# Convert the regular data frame into a dplyr table
tbl_raw_data <- tbl_df(raw_data)
```


```{r}
pryr::object_size(tbl_raw_data)
```



```{r}
dim(tbl_raw_data)

str(tbl_raw_data)
```




```{r,results='asis'}
for (i in colnames(tbl_raw_data)) {
    p <- summary(tbl_raw_data[,i])
    
    cat("<h4>",i,"</h4>")
    cat("<p>","Some stuff here just to make an easy test","</p>")
    cat("<ul>")
    for (j in p) {
        cat("<li>",j,"</li>")}
    cat("</ul>")
    cat("<p>","More text here!!","</p>")
    
    cat("<code>")
    cat(p)
    cat("</code>")
    
    cat("<code>")
    cat(str(tbl_raw_data$LATITUDE))
    cat("</code>")
    
    
}
```



#### 3.2. Output