---
title: '`Course Project 2` Reproducible Research'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
params:
  author: "Anderson H Uyekita"
  course: "Reproducible Research"
  course_url: "https://www.coursera.org/learn/reproducible-research"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  course_start: !r lubridate::ymd("2022/06/24")
  course_finish: !r lubridate::ymd("2022/06/28")
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project-2_reproducible-research'
  project_repo: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2'
  codebook: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
library(shiny)
```


`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s)", "./instructions.md")}else{sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s){target='_blank' rel='noopener'}", params$instructions)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4C4; README: [README.md](%s)", "./README.md")}else{sprintf(fmt = "* &#x1F4C4; README: [README.md](%s){target='_blank' rel='noopener'}", params$project_repo)}`

--------------------------------------------------------------------------------

## Codebook

### 1. Requirements

```{r,echo=TRUE,message=FALSE,warning=FALSE,include=FALSE}
library(tidyverse)
library(pryr)
library(kableExtra)
library(magrittr)
```

### 2. Scripts

Course Project 2 aims to create an Rmd file to create a **Literate Programming**. To do so, I have made the `README.Rmd` which merges text and code.

#### 2.1. README



### 3. Output and Input Details

Course Project 2 has only one input dataset from NOAA and there is no output file.

#### 3.1. Inputs

The file from NOAA should be downloaded from:

* Source: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2
* Download file size: 47 Mbytes

##### 3.1.1. Raw dataset

The content of `repdata_data_StormData.csv.bz2` is a `CSV` file compressed in a `bz2` file.

```{r loading-storm-data,echo=FALSE,cache=TRUE}
# Loading raw data.
raw_data <- read.csv(file = "./data/repdata_data_StormData.csv.bz2")

# Convert the regular data frame into a dplyr table
raw_data <- tbl_df(raw_data)
```

* Object Size in R environment: `r pryr::object_size(raw_data)`
* Number of Observations: `r format(x = dim(raw_data)[1], big.mark = ",")`
* Number of Variables: `r dim(raw_data)[2]`

**Structure**

```{r,echo=FALSE}
str(raw_data)
```

**Variables**

```{r,echo=FALSE}
colnames(raw_data)
```

**Summary**

```{r,echo=FALSE}
summary(raw_data)
```

Some variables have not been imported correctly, which will demand adjustments.

##### 3.1.1. Tidy Dataset

The tidy dataset is a subset of the raw dataset because the Course Project does not require using all variables. Initially, I selected the following variables as part of the tidy dataset. Later some variables will be bundled together and may be dropped.

* `EVTYPE`: Event type;
* `INJURIES`: Quantity of injuries;
* `FATALITIES`: Quantity of fatalities;
* `PROPDMG`: Amount in USD in Damages in properties;
* `CROPDMG`: Amount in USD in Damages in crops;
* `PROPDMGEXP`: The exponential in base 10 used to notate the PROPDMG, and;
* `CROPDMGEXP`: The exponential in base 10 used to notate the CROPDMG.

```{r subset, echo = FALSE,cache = TRUE}
# Subsetting the raw dataset.
df_subset <- select(raw_data,EVTYPE,INJURIES,FATALITIES,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)

# Merging exponential column and value column.
df_subset_2 <- df_subset %>%
    
    # Creating a new column PROPDMGABS
    mutate(PROPDMGABS = case_when(
        PROPDMGEXP %in% c("", "-",0) ~ PROPDMG * 1,
        PROPDMGEXP %in% c(1)         ~ PROPDMG * 10,
        PROPDMGEXP %in% c("H","h",2) ~ PROPDMG * 100,
        PROPDMGEXP %in% c("K","k",3) ~ PROPDMG * 1000,
        PROPDMGEXP %in% c(4)         ~ PROPDMG * 10000,
        PROPDMGEXP %in% c(5)         ~ PROPDMG * 100000,
        PROPDMGEXP %in% c("M","m",6) ~ PROPDMG * 1000000,
        PROPDMGEXP %in% c(7)         ~ PROPDMG * 10000000,
        PROPDMGEXP %in% c(8)         ~ PROPDMG * 100000000,
        PROPDMGEXP %in% c("B","b",9) ~ PROPDMG * 1000000000)) %>%
    
    # Creating a new column CROPDMGABS
    mutate(CROPDMGABS = case_when(
        CROPDMGEXP %in% c("?","", 0)   ~ CROPDMG * 1,
        CROPDMGEXP %in% c("H", "h", 2) ~ CROPDMG * 100,
        CROPDMGEXP %in% c("K", "k", 3) ~ CROPDMG * 1000,
        CROPDMGEXP %in% c("M", "m", 6) ~ CROPDMG * 1000000,
        CROPDMGEXP %in% c("B", "b", 9) ~ CROPDMG * 1000000000))

# Standardizing the EVTYPE variable.
df_subset_3 <- df_subset_2 %>%
    
    # Converting all value to be UPPERCASE.
    mutate(EVTYPE = stringr::str_to_upper(EVTYPE)) %>%
    
    # Rules to aggregate events into a new variable called EVTYPE_2
    mutate(EVTYPE_2 = case_when(
        grepl(pattern = "*HURRICANE*|*TYPHOON*", x = EVTYPE) ~ "HURRICANE",
        grepl(pattern = "*TORNADO*", x = EVTYPE) ~ "TORNADO",
        grepl(pattern = "*THUNDERSTORM WIN*|*TSTM*|*FUNNEL CLOUD*|*FUNNEL*|*THUNDERSTORM*", x = EVTYPE) ~ "THUNDERSTORM WIND",
        grepl(pattern = "*FLOOD*|*FLD*", x = EVTYPE) ~ "FLOOD",
        grepl(pattern = "*HAIL*", x = EVTYPE) ~ "HAIL",
        grepl(pattern = "*HIGH WIND*|*STRONG WIND*|WIND", x = EVTYPE) ~ "HIGH WIND",
        grepl(pattern = "*FOREST FIRE*|*WILD FIRES*|*WILDFIRE*", x = EVTYPE) ~ "FOREST FIRE",
        grepl(pattern = "*LIGHTNING*", x = EVTYPE) ~ "LIGHTNING",
        grepl(pattern = "*HEAT*|*DROUGHT*|*EXCESSIVE HEAT*|*RECORD HIGH*|*WARMTH*|*UNSEASONABLY WARM*|*UNSEASONABLY DRY*|*TEMPERATURE RECORD*|*RECORD TEMPERATURE*|*DRY*", x = EVTYPE) ~ "EXCESSIVE HEAT",
        grepl(pattern = "*WINTER WEATHER*|*FROST/FREEZE*|*EXTREME COLD*|*COLD*|*ICE*|*FREEZE*|*FROST*|*WINTRY MIX*|*LOW TEMPERATURE*", x = EVTYPE) ~ "EXCESSIVE COLD",
        grepl(pattern = "*WINTER STORM*|*SLEET*|*SNOW*|*BLIZZARD*|*FREEZING RAIN*", x = EVTYPE) ~ "ICE STORM",
        grepl(pattern = "*LANDSLIDE*|*MUDSLIDE*|*MUD SLIDE*", x = EVTYPE) ~ "LANDSLIDE",
        grepl(pattern = "*AVALANCHE*", x = EVTYPE) ~ "AVALANCHE",
        grepl(pattern = "*FOG*|*SMOKE*", x = EVTYPE) ~ "FOG",
        grepl(pattern = "*DUST*", x = EVTYPE) ~ "DUST",
        grepl(pattern = "*CURRENT*|*SURF*", x = EVTYPE) ~ "CURRENT",
        grepl(pattern = "*TSUNAMI*", x = EVTYPE) ~ "TSUNAMI",
        grepl(pattern = "*WATERSPOUT*", x = EVTYPE) ~ "WATERSPOUT")) %>%
    
    # Creating a OTHER type of event to aggregate minor events.
    mutate(EVTYPE_2 = case_when(
        is.na(EVTYPE_2) ~ "OTHER",
        !is.na(EVTYPE_2) ~ EVTYPE_2)) %>%
    
    # Dropping EVTYPE because it is not necessary to carry it.
    select(-EVTYPE) %>%
    
    # Renaming the EVTYPE_2 to be the new EVTYPE
    mutate(EVTYPE = EVTYPE_2)

# Creating the tidy dataset.
df_tidy <- df_subset_3 %>%
    
    # Dropping the columns: PROPDMGEXP, PROPDMG, CROPDMGEXP, CROPDMG
    select(-c(PROPDMGEXP, PROPDMG, CROPDMGEXP, CROPDMG)) %>%
    
    # Converting EVTYPE into Factor
    mutate(EVTYPE = factor(EVTYPE_2)) %>%
    
    # Dropping
    select(-EVTYPE_2)
```

After the data manipulation, the tidy dataset will have the following characteristics.

* Object Size in R environment: `r pryr::object_size(df_tidy)`
* Number of Observations: `r format(x = dim(df_tidy)[1], big.mark = ",")`
* Number of Variables: `r dim(df_tidy)[2]`

**Structure**

```{r,echo=FALSE}
str(df_tidy)
```

**Variables**

```{r,echo=FALSE}
colnames(df_tidy)
```

**Summary**

```{r,echo=FALSE}
summary(df_tidy)
```



#### 3.2. Output

Course Project 2 do not have any output.

### 4. Publication

Course Project 2 should be published in [Rpubs](https://rpubs.com/AndersonUyekita/course-project-2_reproducible-research).