---
title: '`Course Project 2` Reproducible Research'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Reproducible Research"
  course_url: "https://www.coursera.org/learn/reproducible-research"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  course_start: !r lubridate::ymd("2022/06/24")
  course_finish: !r lubridate::today()
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project-2_reproducible-research'
  project_repo: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2'
  codebook: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/reproducible-research_course-project-2/blob/master/instructions.md'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`

--------------------------------------------------------------------------------

## Impact of Weather Events on Public Health and Economics

### Sinopsis



--------------------------------------------------------------------------------

This report try to explore the storm database from [U.S. National Oceanic and Atmospheric Administration's (NOAA)](http://www.noaa.gov). This study has a brief analysis of: Economics consequences due to storm, hurricane, and similars and which kind of weather situations has the worse consequence to the population; Population Health due to those extreme weather situations and which disasters has the most impact in the population health. As a results, we realized the tornado is the most harmful and the flood has the greatest economic consequences to the population.

## Loading and Processing the Raw Data

From the [U.S. National Oceanic and Atmospheric Administration's (NOAA)][1] storm database we obtained data of storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage. We obtained the files between the years [1950 and 2011]([2]).

[1]: http://www.noaa.gov
[2]: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2

```{r,cache=TRUE}
# 1. Create a data directory
if(!base::file.exists("data")) {
    base::dir.create("data")
}

# 2. Download files and store it in data directory.
if(!base::file.exists("./data/repdata_data_StormData.csv.bz2")){
    utils::download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2",
                         destfile = "./data/repdata_data_StormData.csv.bz2")
}
```

### Required Packages and Settings 

Requirements necessary to reproduce this analysis:
```{r settings, echo = TRUE, warning=FALSE}
# Loading libraries
library(ggplot2)
library(dplyr)

# Force the R environment to use english settings
Sys.setlocale("LC_ALL","English")
```

### Data Processing (Reading data from 1950 to 2011)

First of all, we read the data from the raw `.csv` file included in the `bz2` archive. The data is a regular delimited file were fileds are delimited with dot comma and missing values are coded as blank fields.

```{r reading, echo = TRUE,cache = TRUE}
raw_data <- read.csv(file = "./data/repdata_data_StormData.csv.bz2")

# Convert the regular data frame into a dplyr table
tbl_raw_data <- tbl_df(raw_data)
```

After reading the bz2 file, the raw data has `r dim(tbl_raw_data)[1]` observations and `r dim(tbl_raw_data)[2]` column (variables), as can be seen by the dim() function.
```{r dimensions, echo = TRUE,cache = TRUE}
dim(tbl_raw_data)
```

The first 6 lines of the raw data set:
```{r head_raw_data, echo = TRUE,cache = TRUE}
head(tbl_raw_data,6)
```

The last 6 lines of the raw data set:
```{r tail_raw_data, echo = TRUE,cache = TRUE}
tail(tbl_raw_data,6)
```


### Questions

1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?
2. Across the United States, which types of events have the greatest economic consequences?

### Subsetting the raw data

This step is necessary to turn the study faster due to the large amount of informations on the raw data. The subset aim to gather only the data necessary to make the desired analysis. Thus, will be select the columns: EVTYPE, INJURIES, FATALITIES, PROPDMG, CROPDMG, PROPDMGEXP, and CROPDMGEXP.
```{r subset, echo = TRUE,cache = TRUE}
dataset <- select(tbl_raw_data,EVTYPE,INJURIES,FATALITIES,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)
```

There are a weird notation to PROPDMG and CROPDMG. Each one need one more column to represent the information of losses (PROPDMGEXP and CROPDMGEXP). The index `H` represent Hundreds, `K` represent Thousand, `M` represent Millions and `B` represent Billions. Thus, those information will be aggregate into one column.
```{r eliminate_exp_column, echo = TRUE}
# The PROPDMGEXP is case sensitive.
dataset <- mutate(dataset,PROPDMGEXP_num = 0 )
dataset <- mutate(dataset,PROPDMGEXP_num = ifelse(test = ((PROPDMGEXP == "H") | (PROPDMGEXP == "h")), yes = 100, no = PROPDMGEXP_num ) )
dataset <- mutate(dataset,PROPDMGEXP_num = ifelse(test = ((PROPDMGEXP == "K") | (PROPDMGEXP == "k")), yes = 1000, no = PROPDMGEXP_num ) )
dataset <- mutate(dataset,PROPDMGEXP_num = ifelse(test = ((PROPDMGEXP == "M") | (PROPDMGEXP == "m")), yes = 1000000, no = PROPDMGEXP_num ) )
dataset <- mutate(dataset,PROPDMGEXP_num = ifelse(test = ((PROPDMGEXP == "B") | (PROPDMGEXP == "b")), yes = 1000000000, no = PROPDMGEXP_num ) )

# Substitution the incomplete information with the compounding of those two column (PROPDMGEXP and PROPDMG).
dataset <- mutate(dataset,PROPDMG = PROPDMGEXP_num*PROPDMG)

# The CROPDMGEXP is case sensitive.
dataset <- mutate(dataset,CROPDMGEXP_num = 0 )
dataset <- mutate(dataset,CROPDMGEXP_num = ifelse(test = ((CROPDMGEXP == "H") | (CROPDMGEXP == "h")), yes = 100, no = CROPDMGEXP_num ) )
dataset <- mutate(dataset,CROPDMGEXP_num = ifelse(test = ((CROPDMGEXP == "K") | (CROPDMGEXP == "k")), yes = 1000, no = CROPDMGEXP_num ) )
dataset <- mutate(dataset,CROPDMGEXP_num = ifelse(test = ((CROPDMGEXP == "M") | (CROPDMGEXP == "m")), yes = 1000000, no = CROPDMGEXP_num ) )
dataset <- mutate(dataset,CROPDMGEXP_num = ifelse(test = ((CROPDMGEXP == "B") | (CROPDMGEXP == "b")), yes = 1000000000, no = CROPDMGEXP_num ) )

# Substitution the incomplete information with the compounding of those two column (CROPDMGEXP and CROPDMG).
dataset <- mutate(dataset,CROPDMG = CROPDMGEXP_num*CROPDMG)

# Subsetting the dataset to show only the necessary column.
tidy_dataset <- select(dataset,EVTYPE,INJURIES,FATALITIES,PROPDMG,CROPDMG)
```

### Results

#### Question 1
* Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

*Answer*

First of all, we need to aggregate `INJURIES` and `FATALITIES` to help us to discern which `EVTYPE` is the most harmful to the population health.
```{r aggregate_by_EVTYPE, echo = TRUE}
# Function aggregate() used to compounding variables.
EVTYPE_harmful <- aggregate( cbind(INJURIES,FATALITIES,INJURIES+FATALITIES) ~ EVTYPE, data = tidy_dataset, FUN = sum)

# Renaming column to keep the dataset organized.
colnames(EVTYPE_harmful) <- c('EVTYPE','INJURIES','FATALITIES','INJURIES_FATALITIES')
```

Assuming a proper register of `EVTYPES` without any typos and any other problem. The most harmful is the TORNADO with 91346 injuries and 5633 fatalities over the dataset period. The table bellow shows the first 6 harmful `EVTYPES`.
```{r sorting, echo = TRUE}
attach(EVTYPE_harmful)
head(EVTYPE_harmful[order(-INJURIES_FATALITIES),],6)
```

Graphically, the most harmful `EVTYPES`.
```{r graf_1, echo = TRUE}
# Subset of first 6 most harmful EVTYPES
bar_plot <- head(EVTYPE_harmful[order(-INJURIES_FATALITIES),],6)

# Barplot using ggplot package
ggplot(bar_plot,aes(x = EVTYPE,y = INJURIES_FATALITIES,group = 1))+
geom_bar(stat = "identity")+
ylab("Sum of Fatalities and Injuries between 1950 - 2011") +
xlab("Event") +
ggtitle ("Total fatalities per event")
```


#### Question 2
2. Across the United States, which types of events have the greatest economic consequences?

*Answer*

First of all, we need to aggregate `PROPDMG` and `CROPDMG` to help us to discern which `EVTYPE` has the the greatest economic consequences.
```{r aggregate_by_EVTYPE2, echo = TRUE}
# Function aggregate() used to compounding variables.
EVTYPE_DMG <- aggregate( cbind(PROPDMG,CROPDMG,PROPDMG + CROPDMG) ~ EVTYPE, data = dataset, FUN = sum)

# Renaming column to keep the dataset organized.
colnames(EVTYPE_DMG) <- c('EVTYPE','PROPDMG','CROPDMG','PROPDMG_CROPDMG')
```


Assuming a proper register of `EVTYPES` without any typos and any other problem. The `EVTYPES` that has the the greatest economic consequences is the `FLOOD` with 150,319,678,250 of dollars over the dataset period. The table bellow shows the results of others `EVTYPES`.
```{r sorting2, echo = TRUE}
attach(EVTYPE_DMG)
head(EVTYPE_DMG[order(-PROPDMG_CROPDMG),],6)
```

Graphically, the `EVTYPES` with the the greatest economic consequences.
```{r graf_2, echo = TRUE}
# Subset of first 6 most harmful EVTYPES
bar_plot2 <- head(EVTYPE_DMG[order(-PROPDMG_CROPDMG),],6)

# Barplot using ggplot package
ggplot(bar_plot2,aes(x = EVTYPE,y = PROPDMG_CROPDMG,group = 1))+
geom_bar(stat = "identity")+
ylab("Sum of Economic Consequences between 1950 - 2011") +
xlab("Event") +
ggtitle ("Total fatalities per event")
```